---
title: "Advanced hooks: useBlocker and useOutletContext"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Advanced hooks: useBlocker and useOutletContext}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Overview

This vignette covers two hooks that address specific routing patterns beyond
basic data loading and navigation:

| Hook | What it does |
|---|---|
| `useBlocker()` | Intercept navigation — e.g. warn about unsaved changes |
| `useOutletContext()` | Read data passed from a parent route through `Outlet(context = ...)` |

Both hooks require a data router: `RouterProvider()` with
`createHashRouter()`, `createBrowserRouter()`, or `createMemoryRouter()`.

## Two ways to inject hook values

Every hook wrapper in **reactRouter** offers two mutually exclusive ways to
turn a hook value into UI. Pick whichever is more ergonomic for the case at
hand — the API is identical across all hooks (`useLoaderData()`,
`useParams()`, `useBlocker()`, `useOutletContext()`, …).

**1. R-idiomatic (`into` + `as` + `selector`)** — the default.
The hook value is cloned into an existing component as a single prop.
Good for displaying one field, populating a data grid's `rows`, or feeding
an input's `value`:

```r
useOutletContext(tags$span(), selector = "user.name")
```

**2. Native render-prop (`render = JS(...)`)** — the escape hatch.
Pass a JS function `(value) => ReactNode`. Mirrors the official React
Router pattern and is the right choice when one prop is not enough — e.g.
combining several fields, conditional rendering, or template strings:

```r
useOutletContext(render = JS("u => `${u.name} (${u.role})`"))
```

When `render` is supplied, `into`, `as`, and `selector` are ignored —
apply any selection inside the JS function directly.

---

## `useBlocker()`

`useBlocker()` intercepts navigation when a JavaScript condition returns
`true`. It exposes a `state` field:

| State | Meaning |
|---|---|
| `"unblocked"` | No navigation is being intercepted (default) |
| `"blocked"` | A navigation was intercepted — show a confirmation UI |
| `"proceeding"` | User confirmed; navigation is in progress |

### Arguments

| Argument | Required | Description |
|---|---|---|
| `into` | yes\* | Component that receives the value |
| `as` | no | Prop to inject into. Defaults to `"children"` |
| `selector` | no | Field from the blocker object. Defaults to `"state"` |
| `render` | no | `JS()` function `(value) => ReactNode`. When supplied, replaces `into`/`as`/`selector`. |
| `shouldBlock` | no | A `JS()` function receiving `{ currentLocation, nextLocation, historyAction }` and returning `true` to block or `false` to allow. Defaults to `FALSE` (never blocks). |

\* `into` is required unless `render` is supplied.

### Show blocker state

```r
useBlocker(
  tags$span(),
  shouldBlock = JS(
    "({ currentLocation, nextLocation }) =>
      currentLocation.pathname !== nextLocation.pathname"
  )
)
# renders "unblocked", "blocked", or "proceeding"
```

### Conditional UI with `render`

Because `selector = "state"` only exposes the state string, use `render` when
you need to branch on the full blocker object (e.g. show the target URL in
the confirmation prompt):

```r
useBlocker(
  shouldBlock = should_block,
  render = JS(
    "b => b.state === 'blocked'
      ? `Leave for ${b.location.pathname}?`
      : null"
  )
)
```

### Full example

The `render` function receives the full blocker object, including `b.proceed()` and
`b.reset()` methods. Use them inside `onClick` handlers to build a real
confirmation dialog:

```{r}
library(reactRouter)
library(htmltools)

should_block <- JS(
  "({ currentLocation, nextLocation }) =>
    currentLocation.pathname !== nextLocation.pathname"
)

blocker_ui <- JS(
  "b => {
    const state = React.createElement('code', null, b.state);
    if (b.state !== 'blocked') return state;
    return React.createElement(React.Fragment, null,
      state, ' — ',
      React.createElement('button', { onClick: () => b.proceed() }, 'Leave'),
      ' ',
      React.createElement('button', { onClick: () => b.reset() }, 'Stay')
    );
  }"
)

RouterProvider(
  router = createHashRouter(
    Route(
      path = "/",
      element = div(
        tags$nav(tags$ul(
          tags$li(NavLink(to = "/",      "Home")),
          tags$li(NavLink(to = "/other", "Other page"))
        )),
        tags$hr(),
        Outlet()
      ),
      Route(
        index = TRUE,
        element = div(
          tags$h3("Home — blocker active"),
          tags$p(
            tags$strong("Blocker state: "),
            useBlocker(shouldBlock = should_block, render = blocker_ui)
          ),
          tags$p(
            style = "color: #555; font-size: 0.9em;",
            "Click 'Other page'. The state changes to ",
            tags$code('"blocked"'), " and two buttons appear.",
            tags$br(),
            tags$em("Leave"), " calls ", tags$code("b.proceed()"),
            " to confirm navigation;",
            tags$br(),
            tags$em("Stay"), " calls ", tags$code("b.reset()"),
            " to cancel and return to ", tags$code('"unblocked"'), "."
          )
        )
      ),
      Route(
        path = "other",
        element = div(
          tags$h3("Other page"),
          tags$p("No blocker here. Navigate freely.")
        )
      )
    )
  )
)
```

### Conditional blocking

Pass any condition as the `shouldBlock` JS function:

```r
# Block only when navigating to a specific path
JS("({ nextLocation }) => nextLocation.pathname === '/confirm'")

# Always block (useful for testing)
JS("() => true")

# Never block
FALSE  # the default
```

---

## `useOutletContext()`

Parent routes can pass data to child routes through `Outlet(context = ...)`.
Any child route reads that data with `useOutletContext()`, avoiding prop-drilling
through deeply nested layouts.

### Arguments

| Argument | Required | Description |
|---|---|---|
| `into` | yes\* | Component that receives the value |
| `as` | no | Prop to inject into. Defaults to `"children"` |
| `selector` | no | Field to extract from the context object. Dotted paths like `"user.name"` navigate nested objects. `NULL` (default) injects the full context. |
| `render` | no | `JS()` function `(context) => ReactNode`. When supplied, replaces `into`/`as`/`selector`. |

\* `into` is required unless `render` is supplied.

### Passing context from a parent route

Add a `context` argument to `Outlet()`:

```r
Route(
  path = "/",
  element = div(
    # Pass a named list — any R object that serialises to JSON
    Outlet(context = list(user = "Alice", role = "admin")),
    NavLink(to = "/profile", "Profile")
  ),
  Route(
    path = "profile",
    element = div(
      tags$p("Name: ",  useOutletContext(tags$span(), selector = "user")),
      tags$p("Role: ",  useOutletContext(tags$span(), selector = "role"))
    )
  )
)
```

### Full example

```{r}
library(reactRouter)
library(htmltools)

current_user <- list(name = "Alice", role = "admin", email = "alice@example.com")

RouterProvider(
  router = createHashRouter(
    Route(
      path = "/",
      element = div(
        tags$nav(tags$ul(
          tags$li(NavLink(to = "/",        "Dashboard")),
          tags$li(NavLink(to = "/profile", "Profile")),
          tags$li(NavLink(to = "/settings","Settings"))
        )),
        tags$hr(),
        Outlet(context = current_user)
      ),
      Route(
        index = TRUE,
        element = div(
          tags$h3("Dashboard"),
          tags$p(
            "Welcome back, ",
            tags$strong(useOutletContext(tags$span(), selector = "name")),
            "!"
          ),
          tags$p("Your role: ", useOutletContext(tags$code(), selector = "role"))
        )
      ),
      Route(
        path = "profile",
        element = div(
          tags$h3("Profile"),
          tags$p("Name: ",  useOutletContext(tags$span(), selector = "name")),
          tags$p("Email: ", useOutletContext(tags$span(), selector = "email")),
          tags$p("Role: ",  useOutletContext(tags$span(), selector = "role")),
          tags$details(
            tags$summary("Full context (JSON)"),
            tags$pre(useOutletContext(tags$span()))
          )
        )
      ),
      Route(
        path = "settings",
        element = div(
          tags$h3("Settings"),
          tags$p(
            "Editing settings for ",
            tags$strong(useOutletContext(tags$span(), selector = "name"))
          )
        )
      )
    )
  )
)
```

The context is passed once at the layout level and is available to every
child route without repeating it in each `element`.

### Combining fields with `render`

`selector` extracts a single field. When you need to compose several fields
into one output — a greeting, a formatted label, a conditional element —
use `render` instead:

```r
useOutletContext(
  render = JS(
    "u => u.role === 'admin'
      ? `${u.name} (admin — ${u.email})`
      : u.name"
  )
)
```

This matches the React Router v7 idiom where the component itself decides
how to consume the hook value, at the cost of a small amount of inline JS.
