---
title: "Security considerations"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Security considerations}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

This vignette walks through the few things to keep in mind when shipping a Shiny app that uses `reactRouter`. You don't need a security background to follow it — the goal is to flag the handful of patterns that matter and the mistakes that are easy to avoid.

## The one idea to remember

`reactRouter` is a wrapper around React Router, and **all of its routing logic runs in the user's browser, not on the Shiny server**. Routes, `loader = JS(...)`, `action = JS(...)`, and any other `JS()` snippet are part of the page that ships to every visitor. They can read it, edit it in the browser's developer tools, and call any URL it calls.

Almost every recommendation below follows from that single fact.

## 1. Don't put secrets in `JS()`

Anything you put inside a `JS()` string is visible to anyone who opens the page. That includes API keys, tokens, and database credentials.

```{r}
library(reactRouter)

# NOT OK — the API key is shipped to every visitor
Route(
  path = "/private",
  loader = JS("async () => {
    const r = await fetch('https://api.example.com/data', {
      headers: { Authorization: 'Bearer sk-LIVE-1234...' }
    });
    return r.json();
  }"),
  element = useLoaderData(tags$pre())
)
```

If a loader needs authenticated data, fetch it from your own Shiny server (or an API you control) and let the server attach the credential.

## 2. Authentication belongs to Shiny, not the router

A loader like `if (!isLoggedIn()) return redirect("/login")` is convenient, but it isn't a real lock — a curious user can just delete the check from their browser and reach the route.

For real protection, do the check where the user can't tamper with it:

* **In front of Shiny** — Posit Connect, Shinyapps.io, ShinyProxy, or a reverse proxy that requires a login before the app loads.
* **Inside the Shiny server** — packages like [shinymanager](https://datastorm-open.github.io/shinymanager/), or your own checks against `session$user`.

`reactRouter` is a good fit for *showing* a login screen or a "you're not allowed to see this" page, but the decision about whether the user is allowed to see the data should be made on the server.

A useful test: if you removed every `JS()` check from your app, would the user see anything they shouldn't? If yes, the check is in the wrong place.

## 3. React escapes text — keep it that way

When you display loader or action data with `useLoaderData()` / `useActionData()`, React renders it as plain text. Even if a value contains HTML or `<script>` tags, it shows up as literal characters, not as code. You get this protection for free.

```{r}
# Safe: this renders as the literal text "<img src=x onerror=alert(1)>"
Route(
  path = "/profile",
  loader = JS("async () => ({ bio: '<img src=x onerror=alert(1)>' })"),
  element = useLoaderData(tags$div(), selector = "bio")
)
```

The only way to lose this protection is to opt out of it explicitly with `dangerouslySetInnerHTML` in custom JavaScript. Don't, unless you've sanitised the input first (e.g. with [DOMPurify](https://github.com/cure53/DOMPurify)).

## 4. Encode URL parameters in `fetch()` calls

If a loader takes a path parameter (like `:id`) and pastes it into a URL, wrap it in `encodeURIComponent()`. Otherwise a crafted link can append extra query parameters or path segments to the upstream request.

```{r}
# Vulnerable
loader_bad <- JS("async ({ params }) => {
  const r = await fetch(`https://api.example.com/films/${params.id}`);
  return r.json();
}")

# Safe
loader_good <- JS("async ({ params }) => {
  const id = encodeURIComponent(params.id);
  const r  = await fetch(`https://api.example.com/films/${id}`);
  return r.json();
}")
```

The same applies to values read from `useSearchParams()` before they go into a URL.

## 5. Validate redirect targets

`redirect()` blocks dangerous URL schemes such as `javascript:` and `data:`,
and protocol-relative targets like `//evil.example.com/path` (which the
browser would otherwise interpret as an absolute URL on another origin):

```{r}
redirect("javascript:alert(1)")
#> Error: redirect(): refusing unsafe URL scheme in `to` = "javascript:alert(1)".
redirect("//evil.example.com/path")
#> Error: redirect(): refusing protocol-relative URL in `to` = "//evil.example.com/path". ...
```

If you genuinely need a cross-origin redirect, spell out the full
`https://...` URL. And if `to` ever comes from user input, still don't
rely on the built-in checks alone -- match it against an allowlist.

But if you build a redirect target from user input (for example a `?next=` parameter after a sign-in), it's safer to compare against a list of allowed paths than to try to filter the string yourself:

```{r}
nextParam  <- "/dashboard"   # in real code, from useSearchParams()
allowedNext <- c("/", "/dashboard", "/profile", "/reports")

target <- if (nextParam %in% allowedNext) nextParam else "/"
Route(path = "/post-login", loader = redirect(target), element = NULL)
```

## 6. A note on `Form` and `action`

`reactRouter` exposes a `Form` component and route `action`s for completeness, but they also run in the browser. If you use them to save data, the actual write needs to happen on a server endpoint you control — that's where authentication, validation, and storage belong. For most Shiny apps, sticking to regular Shiny inputs and `observeEvent()` on the server is simpler and safer.

## 7. Content-Security-Policy (optional, but a good idea)

If you're hosting your app yourself and want an extra layer of defence, add a `Content-Security-Policy` header. Among other things, it lets you list exactly which external APIs your app is allowed to talk to, so a bug elsewhere in your code can't quietly send data to a third party. A starting point for a Shiny app:

```
Content-Security-Policy:
  default-src 'self';
  script-src  'self' 'unsafe-inline';
  style-src   'self' 'unsafe-inline';
  connect-src 'self' https://api.example.com;
  img-src     'self' data: https:;
  frame-ancestors 'none';
```

`'unsafe-inline'` for scripts is currently needed by **Shiny** itself (its
client uses inline scripts to bootstrap the session); `reactRouter` does not
require it. If a future Shiny release tightens this, the policy above can
drop `'unsafe-inline'`.

## Reporting an issue

If you find a security problem in `reactRouter`, please email the maintainer of this R package rather than opening a public issue, so a fix can be released before the details are widely shared.
