---
title: "Using Form, useFetcher, useFetchers, and useRevalidator"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using Form, useFetcher, useFetchers, and useRevalidator}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Overview

React Router v7 provides hooks and components for managing data fetching and
navigation state:

| Function | Role |
|---|---|
| `Form()` | A `<form>` that triggers a route loader (GET) or action (POST) |
| `useFetcher()` | Read a fetcher's `state` and `data` |
| `useFetchers()` | Read the states of all active fetchers |
| `useRevalidator()` | Re-run the current route's loader; read revalidation `state` |

These only work inside a data router
(`RouterProvider()`, `createHashRouter()`, or `createMemoryRouter()`).

---

## `Form()` with GET — search and `useLoaderData`

A GET `Form()` updates the URL's search params and triggers the route's
loader. Combine it with `useLoaderData()` to display results and
`useSearchParams()` to reflect the current query.

```{r}
library(reactRouter)
library(htmltools)

RouterProvider(
  router = createHashRouter(
    Route(
      path = "/",
      element = div(
        tags$h2("Search Example"),
        Form(
          method = "get",
          action = "/search",
          tags$input(type = "search", name = "q", placeholder = "Search..."),
          tags$button(type = "submit", "Search")
        ),
        Outlet()
      ),
      Route(index = TRUE, element = div(tags$p("Enter a query above."))),
      Route(
        path = "search",
        loader = JS("({ request }) => {
          const q = new URL(request.url).searchParams.get('q') || '';
          const db = ['Alice', 'Bob', 'Charlie', 'David', 'Eve'];
          const hits = db.filter(n => n.toLowerCase().includes(q.toLowerCase()));
          return { count: hits.length, results: hits.join(', '), query: q };
        }"),
        element = div(
          tags$p("Query: ",   useSearchParams(tags$strong(), param = "q")),
          tags$p("Results: ", useLoaderData(tags$strong(),   selector = "count")),
          tags$pre(           useLoaderData(tags$span(),     selector = "results"))
        )
      )
    )
  )
)
```

### Navigation state while loading

Use `useNavigation()` in the parent layout to show `"idle"` / `"loading"` /
`"submitting"` feedback while the loader runs. The snippet below extends the
search example above by adding a status line to the root layout:

```{r}
RouterProvider(
  router = createHashRouter(
    Route(
      path = "/",
      element = div(
        tags$h2("Search with Navigation State"),
        tags$p("Status: ", useNavigation(tags$strong(), selector = "state")),
        Form(
          method = "get",
          action = "/search",
          tags$input(type = "search", name = "q", placeholder = "Search..."),
          tags$button(type = "submit", "Search")
        ),
        Outlet()
      ),
      Route(index = TRUE, element = div(tags$p("Enter a query above."))),
      Route(
        path = "search",
        loader = JS("async ({ request }) => {
          await new Promise(r => setTimeout(r, 800));
          const q = new URL(request.url).searchParams.get('q') || '';
          const db = ['Alice', 'Bob', 'Charlie', 'David', 'Eve'];
          const hits = db.filter(n => n.toLowerCase().includes(q.toLowerCase()));
          return { count: hits.length, results: hits.join(', ') };
        }"),
        element = div(
          tags$p("Query: ",   useSearchParams(tags$strong(), param = "q")),
          tags$p("Results: ", useLoaderData(tags$strong(),   selector = "count")),
          tags$pre(           useLoaderData(tags$span(),     selector = "results"))
        )
      )
    )
  )
)
```

---

## `Form()` with POST — action and `useActionData`

A POST `Form()` triggers the route's `action`. The action result is available
via `useActionData()`.

```{r}
RouterProvider(
  router = createHashRouter(
    Route(
      path = "/",
      action = JS("async ({ request }) => {
        const fd     = await request.formData();
        const metric = fd.get('metric');
        const from   = fd.get('from');
        const to     = fd.get('to');
        return { metric, period: from + ' to ' + to };
      }"),
      element = div(
        tags$h2("Report Generator"),
        Form(
          method = "post",
          tags$select(name = "metric",
            tags$option(value = "Revenue",    "Revenue"),
            tags$option(value = "Page Views", "Page Views")
          ),
          tags$input(type = "date", name = "from"),
          tags$input(type = "date", name = "to"),
          tags$button(type = "submit", "Generate")
        ),
        tags$p("Metric: ", useActionData(into = tags$span(), selector = "metric")),
        tags$p("Period: ", useActionData(into = tags$span(), selector = "period"))
      )
    )
  )
)
```

---

## `useFetcher()`

`useFetcher()` reads the state of a fetcher instance. It is useful when you
need to observe the loading or submission state independently — e.g. to show a
spinner next to an inline element without affecting the global navigation state.

Use `Form(navigate = FALSE, fetcherKey = "...")` to submit data without a
page-level navigation. Then use `useFetcher()` with the same key to read that
fetcher's state and data.

### Arguments

| Argument | Required | Description |
|----------|----------|-------------|
| `into` | yes\* | Component that receives the value |
| `as` | no | Prop to inject into. Defaults to `"children"` |
| `selector` | no | Field from the fetcher object. Defaults to `"state"` |
| `fetcherKey` | no | Key to identify a specific fetcher instance |

\* `into` is required unless `render` is supplied.

The fetcher object fields available via `selector`:

| Selector | Value |
|---|---|
| `"state"` | `"idle"` / `"loading"` / `"submitting"` |
| `"data"` | Full response from the loader or action |
| `"data.myField"` | Nested field from the response |

### Full example: inline search without navigation

```{r}
library(reactRouter)
library(htmltools)
library(jsonlite)

people_json <- jsonlite::toJSON(
  data.frame(name = c("Alice", "Bob", "Charlie", "David", "Eve")),
  dataframe = "rows",
  auto_unbox = TRUE
)

RouterProvider(
  router = createHashRouter(
    Route(
      path = "/",
      element = div(
        tags$h2("useFetcher Example — Inline Search"),
        tags$p(
          "Type a name and click Search. The URL does not change.",
          style = "color: #555;"
        ),
        Form(
          navigate  = FALSE,
          fetcherKey = "search",
          method    = "get",
          action    = "/search",
          style     = "display: flex; gap: 8px; margin-bottom: 16px;",
          tags$input(
            type = "search", name = "q", placeholder = "Search names...",
            style = "padding: 6px; font-size: 14px;"
          ),
          tags$button(type = "submit", "Search", style = "padding: 6px 12px; cursor: pointer;")
        ),
        tags$p(
          tags$strong("Fetcher state: "),
          useFetcher(tags$span(), fetcherKey = "search", selector = "state")
        ),
        tags$p(
          tags$strong("Matches found: "),
          useFetcher(tags$span(), fetcherKey = "search", selector = "data.count")
        ),
        tags$details(
          tags$summary("Raw results (JSON)"),
          tags$pre(useFetcher(tags$span(), fetcherKey = "search", selector = "data.results"))
        ),
        tags$hr(),
        Outlet()
      ),
      Route(index = TRUE, element = div(tags$p("Enter a name above and click Search."))),
      Route(
        path = "search",
        loader = JS(sprintf(
          "async ({ request }) => {
            await new Promise(resolve => setTimeout(resolve, 800));
            const q = new URL(request.url).searchParams.get('q') || '';
            const db = %s;
            const hits = db.filter(r => r.name.toLowerCase().includes(q.toLowerCase()));
            return { count: hits.length, results: hits };
          }",
          people_json
        )),
        element = div()
      )
    )
  )
)
```

---

## `useFetchers()`

`useFetchers()` returns all currently active fetchers. This is useful for a
**global loading indicator** that shows whenever any background fetch is in
progress, regardless of which component started it.

### Arguments

| Argument | Required | Description |
|----------|----------|-------------|
| `into` | yes\* | Component that receives the value |
| `as` | no | Prop to inject into. Defaults to `"children"` |
| `selector` | no | Field to extract from **each** fetcher. If `NULL` (default), returns the full array as JSON |

### Full example: monitor all active fetchers

```{r}
library(reactRouter)
library(htmltools)
library(jsonlite)

people_json <- jsonlite::toJSON(
  data.frame(name = c("Alice", "Bob", "Charlie", "David", "Eve")),
  dataframe = "rows",
  auto_unbox = TRUE
)

loader_js <- JS(sprintf(
  "async ({ request }) => {
    await new Promise(r => setTimeout(r, 1000));
    const q = new URL(request.url).searchParams.get('q') || '';
    const db = %s;
    return db.filter(r => r.name.toLowerCase().includes(q.toLowerCase()));
  }",
  people_json
))

RouterProvider(
  router = createHashRouter(
    Route(
      path = "/",
      element = div(
        tags$h2("useFetchers Example — Global Loading Indicator"),
        tags$p(
          "Submit one or more search forms. Each uses a different fetcher key.",
          style = "color: #555;"
        ),
        tags$p(
          tags$strong("All fetcher states: "),
          useFetchers(tags$code(), selector = "state")
        ),
        tags$hr(),

        tags$h3("Search A"),
        Form(
          navigate = FALSE, fetcherKey = "search-a",
          method = "get", action = "/search",
          style = "display: flex; gap: 8px; margin-bottom: 8px;",
          tags$input(type = "search", name = "q", placeholder = "Name...", style = "padding: 4px;"),
          tags$button(type = "submit", "Search A", style = "cursor: pointer;")
        ),
        tags$p("State A: ",   useFetcher(tags$code(), fetcherKey = "search-a", selector = "state")),
        tags$p("Results A: ", useFetcher(tags$span(), fetcherKey = "search-a", selector = "data")),

        tags$h3("Search B"),
        Form(
          navigate = FALSE, fetcherKey = "search-b",
          method = "get", action = "/search",
          style = "display: flex; gap: 8px; margin-bottom: 8px;",
          tags$input(type = "search", name = "q", placeholder = "Name...", style = "padding: 4px;"),
          tags$button(type = "submit", "Search B", style = "cursor: pointer;")
        ),
        tags$p("State B: ",   useFetcher(tags$code(), fetcherKey = "search-b", selector = "state")),
        tags$p("Results B: ", useFetcher(tags$span(), fetcherKey = "search-b", selector = "data")),

        tags$hr(),
        Outlet()
      ),
      Route(index = TRUE, element = div()),
      Route(path = "search", loader = loader_js, element = div())
    )
  )
)
```

When no fetchers are active, `useFetchers(selector = "state")` renders `"[]"`.

---

## `useRevalidator()` — re-run the loader

Revalidation re-runs the current route's loader **without a page-level
navigation**. This is distinct from navigation (which changes the URL) and
from fetchers (which load a different route in the background).

To trigger revalidation from a button, inject the hook's `revalidate` function
as the `onClick` handler using `as = "onClick"` and `selector = "revalidate"`:

```r
useRevalidator(
  tags$button("Refresh", style = "padding: 6px 14px; cursor: pointer;"),
  as       = "onClick",
  selector = "revalidate"
)
```

### `useRevalidator()` arguments

| Argument | Required | Description |
|----------|----------|-------------|
| `into` | yes\* | Component that receives the value |
| `as` | no | Prop to inject into. Defaults to `"children"` (shows state string) |
| `selector` | no | Field from the revalidator. Defaults to `"state"` |

\* `into` is required unless `render` is supplied.

Use `selector = "state"` to read `"idle"` / `"loading"`, and `selector = "revalidate"` to get the function that triggers revalidation.

### Full example: manual refresh button

```{r}
library(reactRouter)
library(htmltools)

RouterProvider(
  router = createHashRouter(
    Route(
      path = "/",
      loader = JS(
        "async () => {
          await new Promise(r => setTimeout(r, 1000));
          return { time: new Date().toLocaleTimeString() };
        }"
      ),
      element = div(
        tags$h2("useRevalidator Example"),
        tags$p(
          "Click the button to re-run the loader and refresh the timestamp.",
          style = "color: #555;"
        ),
        tags$p(
          tags$strong("Loaded at: "),
          useLoaderData(tags$span(), selector = "time")
        ),
        tags$p(
          tags$strong("Revalidation state: "),
          useRevalidator(tags$span(), selector = "state")
        ),
        useRevalidator(
          tags$button(
            "Refresh (revalidate loader)",
            style = "padding: 6px 14px; cursor: pointer;"
          ),
          as       = "onClick",
          selector = "revalidate"
        ),
        tags$hr(),
        Outlet()
      ),
      Route(index = TRUE, element = div(tags$p("Home page.")))
    )
  )
)
```

---

## Comparison

| Hook / Component | Triggers | State visible via | URL changes? |
|---|---|---|---|
| `Form()` GET + `useLoaderData()` | page navigation | `useNavigation(selector = "state")` | Yes |
| `Form()` POST + `useActionData()` | page navigation | `useNavigation(selector = "state")` | No |
| `Form(navigate = FALSE)` + `useFetcher()` | background fetch | `useFetcher(selector = "state")` | No |
| `useRevalidator()` | loader re-run | `useRevalidator(selector = "state")` | No |
