---
title: "Using useLoaderData"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using useLoaderData}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Overview

`useLoaderData()` connects React Router's
[`useLoaderData()`](https://reactrouter.com/6.30.0/hooks/use-loader-data) hook
to your UI. It calls the hook and injects the result into a component — either
as text content inside a plain HTML element, or as a structured JavaScript value
into any prop of a `shiny.react`-based component.

Use it inside a `Route` that has a `loader` function.

## Arguments

| Argument | Required | Description |
|----------|----------|-------------|
| `into` | yes | The component that will receive the loader data. Can be a plain `htmltools` tag (e.g. `tags$h3()`) or a `shiny.react`-based component. |
| `as` | no | The prop name to inject data into. Defaults to `"children"`. |
| `selector` | no | A key to extract from the loader data object. If `NULL` (default), the entire loader data is passed. |

## Rendering text in HTML elements

Pass any `htmltools` tag as `into` to render loader data as text. This is the
simplest use case — no extra configuration needed:

```r
# Render a single field as text
useLoaderData(tags$h3(), selector = "name")      # renders the "name" field
useLoaderData(tags$pre())                         # renders full JSON string
```

Full example with dynamic route parameters:

```{r}
library(reactRouter)
library(htmltools)
library(dplyr)

people_json <- jsonlite::toJSON(dplyr::starwars, dataframe = "rows", auto_unbox = TRUE)

RouterProvider(
  router = createHashRouter(
    Route(
      path = "/",
      element = tags$div(
        NavLink(to = "/", "Home"), " | ",
        NavLink(to = "/people/1", "Luke"), " | ",
        NavLink(to = "/people/2", "C-3PO"), " | ",
        NavLink(to = "/people/3", "R2-D2"),
        tags$hr(),
        Outlet()
      ),
      Route(index = TRUE, element = p("Select a person.")),
      Route(
        path = "people/:id",
        loader = JS(sprintf(
          "({ params }) => {
            const db = %s;
            const person = db[params.id - 1];
            if (!person) throw new Response('Not found', { status: 404 });
            return person;
          }",
          people_json
        )),
        element = div(
          useLoaderData(into = tags$h3(), selector = "name"),
          useLoaderData(into = tags$p(),  selector = "gender")
        ),
        errorElement = div(tags$p("Not found."), NavLink(to = "/", "Back"))
      )
    )
  )
)
```

## Injecting structured data into shiny.react components

Pass a `shiny.react`-based component as `into` and specify the target prop with
`as`. The loader data (or a `selector` field from it) is injected as the raw
JavaScript value — not converted to a string.

This is needed whenever a component expects an actual array or object for a
prop, for example `rows` in a data grid or `options` in an autocomplete.

```{r}
library(reactRouter)
library(muiDataGrid)
library(htmltools)

loader <- JS("async () => {
  const res = await fetch('https://swapi.info/api/people');
  return (await res.json()).map((p, i) => ({
    id: i, name: p.name, height: p.height
  }));
}")

RouterProvider(
  router = createHashRouter(
    Route(
      path = "/",
      loader = loader,
      element = div(
        style = "height: 400px;",
        useLoaderData(
          muiDataGrid::DataGrid(
            columns = list(
              list(field = "name", headerName = "Name", flex = 1),
              list(field = "height", headerName = "Height", width = 100)
            )
          ),
          as = "rows"
        )
      )
    )
  )
)
```

When the loader returns an object with multiple fields, use `selector` to
extract the relevant one:

```{r}
# -- JS loader: fetches Star Wars characters from SWAPI, filters by ?name= param ----
loader_people <- JS(
  "
  async ({ request }) => {
    const filter = new URL(request.url).searchParams.getAll('name');
    const data = await fetch('https://swapi.info/api/people').then(r => r.json());
    return {
      names: data.map(p => p.name).sort(),
      people: filter.length ? data.filter(p => filter.includes(p.name)) : data
    };
  }
"
)

# -- UI --------------------------------------------------------------------
RouterProvider(
  router = createHashRouter(
    Route(
      path = "/",
      loader = loader_people,
      element = div(
        h2("Star Wars Characters"),
        p(
          "Select characters to filter the table, or clear the selection to show all."
        ),
        useLoaderData(
          into = muiMaterial::Autocomplete(
            onChange = JS(
              "(event, value) => { window.location.hash = value.length ? '/?' + value.map(v => 'name=' + encodeURIComponent(v)).join('&') : '/'; }"
            ),
            renderInput = JS(
              "(params) => React.createElement(window.jsmodule['@mui/material'].TextField, {...params, label: 'Name'})"
            ),
            multiple = TRUE,
            sx = list(width = 300, marginBottom = 2)
          ),
          as = "options",
          selector = "names"
        ),
        useLoaderData(
          into = muiDataGrid::DataGrid(
            columns = list(
              list(field = 'name', headerName = 'Name', flex = 1),
              list(field = 'height', headerName = 'Height', width = 100),
              list(field = 'mass', headerName = 'Mass', width = 100),
              list(field = 'gender', headerName = 'Gender', width = 120),
              list(field = 'birth_year', headerName = 'Birth Year', width = 120)
            ),
            getRowId = JS("function(row) { return(row.name) }"),
            initialState = list(
              pagination = list(paginationModel = list(pageSize = 5))
            ),
            showToolbar = TRUE
          ),
          as = "rows",
          selector = "people"
        )
      )
    )
  )
)
```

## Designing your loader

The React Router community recommends that each loader returns **exactly the
shape the consuming component needs**. This avoids the need for `selector`:

When a single route renders multiple components with different data needs,
consider splitting into nested routes — each with its own loader:

```{r}
library(reactRouter)
library(muiDataGrid)
library(htmltools)

stats_loader <- JS("async () => {
  const res = await fetch('https://swapi.info/api/people');
  const people = await res.json();
  return { summary: people.length + ' characters found' };
}")

people_loader <- JS("async () => {
  const res = await fetch('https://swapi.info/api/people');
  return (await res.json()).map((p, i) => ({
    id: i, name: p.name, height: p.height
  }));
}")

RouterProvider(
  router = createHashRouter(
    Route(
      path = "/",
      loader = stats_loader,
      element = div(
        useLoaderData(tags$h2(), selector = "summary"),
        Outlet()
      ),
      Route(
        index = TRUE,
        loader = people_loader,
        element = div(
          style = "height: 400px;",
          useLoaderData(
            into = muiDataGrid::DataGrid(
              columns = list(
                list(field = "name", "headerName" = "Name", flex = 1),
                list(field = "height", "headerName" = "Height", width = 100)
              )
            ), 
          as = "rows")
        )
      )
    )
  )
)
```

## Works with any shiny.react component

`useLoaderData()` works with any component built on `shiny.react` — it simply
injects a prop value via `React.cloneElement()`. Examples of compatible
packages:

- [`muiMaterial`](https://github.com/lgnbhl/muiMaterial) — MUI Material UI components
- [`muiDataGrid`](https://github.com/lgnbhl/muiDataGrid) — MUI Data Grid
- [`muiCharts`](https://github.com/lgnbhl/muiCharts) — MUI Charts
- [`muiTreeView`](https://github.com/lgnbhl/muiTreeView) — MUI Tree View
- [`shiny.fluent`](https://appsilon.github.io/shiny.fluent/) — Microsoft Fluent UI components
- [`shiny.blueprint`](https://appsilon.github.io/shiny.blueprint/) — Palantir Blueprint UI components

Standard `htmltools` tags (`tags$div`, `tags$h3`, etc.) only support text
injection (the default `as = "children"` with no structured data).

**muiCharts::BarChart** — inject loader data as `dataset`:

```{r}
library(reactRouter)
library(muiCharts)
library(htmltools)

loader <- JS("async () => {
  const res = await fetch('https://swapi.info/api/people');
  const people = await res.json();
  return people.slice(0, 10).map(p => ({
    name: p.name,
    height: Number(p.height) || 0,
    mass: Number(p.mass) || 0
  }));
}")

RouterProvider(
  router = createHashRouter(
    Route(
      path = "/",
      loader = loader,
      element = useLoaderData(
        muiCharts::BarChart(
          series = list(
            list(dataKey = "height", label = "Height (cm)"),
            list(dataKey = "mass",   label = "Mass (kg)")
          ),
          xAxis = list(list(scaleType = "band", dataKey = "name")),
          height = 400
        ),
        as = "dataset"
      )
    )
  )
)
```

## Injecting into multiple props

To inject loader data into more than one prop of the same component, nest
`useLoaderData()` calls:

```{r}
library(reactRouter)
library(muiDataGrid)
library(htmltools)

loader <- JS("async () => {
  const res = await fetch('https://swapi.info/api/people');
  const people = await res.json();
  return {
    columnDefs: [
      { field: 'name',   headerName: 'Name',   flex: 1 },
      { field: 'height', headerName: 'Height', width: 100 }
    ],
    people: people.map((p, i) => ({ id: i, name: p.name, height: p.height }))
  };
}")

RouterProvider(
  router = createHashRouter(
    Route(
      path = "/",
      loader = loader,
      element = div(
        style = "height: 400px;",
        useLoaderData(
          useLoaderData(
            muiDataGrid::DataGrid(),
            as = "rows",
            selector = "people"
          ),
          as = "columns",
          selector = "columnDefs"
        )
      )
    )
  )
)
```

Each call extracts its own `selector` from the same `useLoaderData()` result
and injects it into the specified prop.

## Requirements

`useLoaderData()` requires a **data router** (`createHashRouter`,
`createBrowserRouter`, or `createMemoryRouter`) because only data routers
support `loader` functions. It will not work with the component-based routers
(`HashRouter`, `MemoryRouter`).

The child component's JavaScript bundle must be loaded independently. For
example, `muiDataGrid::DataGrid()` automatically attaches its own HTML
dependency — `reactRouter` does not bundle it.
